---
title: "Improving the Transparency of Power Analyses Through standardised Reporting"
author:
  - name: Cristian Mesquida
    orcid: 0000-0002-1542-8355
    email: c.mesquida.caldentey@tue.nl
    affiliations:
      - name: Eindhoven University of Technology
        department: Industrial Engineering and Innovation Sciences
        address: Den Dolech 2, 5600 MB, Eindhoven
        city: Eindhoven
    roles:
     - conceptualization
     - methodology
     - formal analysis
     - data curation
     - writing — original draft
  - name: Daniël Lakens
    corresponding: true
    orcid: 0000-0002-8393-5316
    email: D.Lakens@tue.nl
    affiliations:
      - name: Eindhoven University of Technology
        department: Industrial Engineering and Innovation Sciences
        address: Den Dolech 2, 5600 MB, Eindhoven
        city: Eindhoven
    roles:
     - conceptualization
     - methodology
     - formal analysis
     - data curation
     - writing — review and editing
  - name: Lisa DeBruine
    orcid: 0000-0002-7523-5539
    email: lisa.debruine@glasgow.ac.uk
    affiliations:
      - name: University of Glasgow
        department: School of Psychology & Neuroscience
    roles:
     - conceptualization
     - methodology
     - formal analysis
     - data curation
     - writing — review and editing
keywords: [Peer Review, Quality Control, Automation, Standardisation, Power Analysis]
author-note:
  disclosures:
    conflict-of-interest: "Authors declare no conflicts of interest."
    data-sharing: "The data and analysis scripts related to this study are publicly available on Zenodo and can be found at XXX."
bibliography: [references.bib, package-citations.bib]
floatsintext: true # Place figures and tables within the text
format:
  apaquarto-html:
    embed-resources: false
    toc: true
    toc-depth: 3
    toc-expand: true
    css: style.css
  apaquarto-docx: default
execute:
  echo: false
abstract: ""
---

<!--
Q: Is metacheck a COI?
-->


```{r}
#| label: setup
#| eval: FALSE

# install the apaquarto extension
quarto::quarto_add_extension("wjschne/apaquarto", no_prompt = TRUE)

# install the main version of Metacheck
pak::pkg_install("scienceverse/metacheck@main")
```

```{r}
#| label: libraries
#| message: FALSE
#| output: FALSE

# Libraries to be loaded
library(metacheck)
library(dplyr)
library(tidyr)
library(flextable)
library(pwr)

# Create BIB file for R packages
knitr::write_bib(
  .packages(),
  file = "package-citations.bib"
)
```

```{r}
#| label: data

# Load RDS object containing articles from 2014 to 2026
psych_science <- readRDS("data/psychsci.rds")
```

## Introduction

:::{.callout-important}
I would start more generically about power analysis importance and prevalence. Sample size is only one use.

@vankelecom_2025 

https://doi.org/10.5334/pb.1318

> ...we conducted a systematic review of 903 published empirical articles across four APA-disciplines, comparing 453 papers published in 2015–2016, with 450 papers from 2020–2021. Although the prevalence of power analysis across different domains in psychology has increased over time (from 9.5% to 30%), it remains insufficient overall. 

:::

An important step when designing a study is to justify the sample size that will be collected. The key aim of a sample size justification is to explain how the collected data are expected to provide valuable information given the inferential goals of the researcher [@lakens2022]. One common sample size justification when the inferential goal is to test a hypothesis is a power analysis, which makes explicit which effect size researchers are interested in detecting. This information is essential both when interpreting the informational value of a study and when interpreting the test results. A study that has high power only for unrealistically large effect sizes is unlikely to contribute useful insights to the published literature. If a hypothesis test yields a non-significant result, the range of effect sizes that could have been detected with low error rates is made explicit by the power analysis, which provides important context to evaluate the test results. For these reasons, researchers need to report all parameters used to compute the power analysis.

:::{.callout-important}
Existing Guildelines:

* https://casrai.org/guides/power-analysis-sample-size-calculation
* https://ori.hhs.gov/education/products/niu_authorship/mistakes/06mistake-a.htm
* https://www.consort-spirit.org/item16a-samplesize

Maybe describe this as no specific reporting guidelines that allow for shorthand consistent representation of essential parameters
:::

Despite the important role power analyses play in hypothesis testing research, there are no reporting guidelines for power analyses. Furthermore, researchers often omit important information [@thibault_2024; @bakker_2020; @mesquida_2025]. For example, Bakker et al [-@bakker_2020] investigated the transparency and reproducibility of power analyses reported in 210 preregistrations submitted to the Open Science Framework and institutional review boards proposals. In 52 (30%) of the reported power analyses, the effect size metric (e.g., Cohen's $d$, $\eta_p^2$) was not specified and in 26 (15%) the effect size magnitude was not specified. Only 34 (20%) power analyses provided enough information to reproduce the power analysis and only 62% of these power analyses pertained to the main hypothesis test in the pre-registration.

Similarly, Thibault et al [-@thibault_2024] assessed the transparency and reproducibility of 95 power analyses reported in published articles. They could only reproduce 2 (2%) without making any assumptions and could likely reproduce another 27 (28%) after making assumptions. For 46 (48%) power analyses, they could not determine whether the power analysis was for a statistical test that appeared in the results section of the publication. Another recurring issue was that power analyses are often based on default options offered by G\*Power (i.e., assuming a medium effect size, $\alpha$ = .05, and power = .80) that may not be relevant to the analysis.

Finally, researchers often fail to justify assumed parameters, such as effect size. For example, Mesquida et al [-@mesquida_2025] reported that in a sample of 144 articles reporting an a priori power analysis, 48 (33%) did not justify the selected effect size. This lack of transparency prevents editors, reviewers, and readers from being able to reproduce power analyses or critically evaluate their validity.

:::{.callout-important}
Maybe give some examples of justifications used?
:::

```{r}
#| label: res_power
#| eval: FALSE ## Set to TRUE to rerun module

# Opt-out LLM
llm_use(FALSE)

# Run the module
res_power <- module_run(psych_science, "power")

# Save results
save(res_power, file = "data/res_power.RData")
```

```{r}
#| label: power_type

# Load data
load("data/res_power.RData")

# Calculate number and proportion of power analyses by type
power_type <- res_power$table |> 
  count(power_type) |> 
  mutate(
    Percent = paste0(
      round((n / sum(n)) * 100, 0),
      "%"
    )
  ) |> 
  select(
    Power = power_type, 
    Cases = n, 
    Percent
    )
```

<!-- 
Before we introduce a module to tell you to report all the essential parts of a power analysis, I think we need to briefly introduce what those parts are
-->

## Components

The report of a power analysis must contain several components in order to be interpretable and reproducible. The reader must be able to determine exactly which test(s) the power analysis corresponds to, and which effect if the model estimates more than one effect. The target of power analysis should be clear: is it calculating power, sample size, effect size, and/or the alpha criterion? The power should be reported explicitly, and not assumed to be a default. The total sample size should be reported explicitly, and where applicable, the sample size per group. The effect size should be specified with a clear metric (e.g., 0.5 Cohen's $d$, not "medium") and justified if it is not the output of the power analysis. Also include the criteria used to determine significance (e.g., the critical alpha and any corrections applied). The assumed parameters of the statistical test must be clear. For example, specifying Student's vs Welch's t-test or the variance components of a linear mixed effects model. Finally, the software used to conduct the power analysis should be cited, including the version, add-on packages used, and ideally with reference to supplemental materials that contain the exact code or specifications if this is too complex to include in the text. 


<!--
## Structured schema

To formalize these reporting requirements and facilitate power analyses to be machine-readable, we have developed a JSON schema that defines the parameters to be reported. The schema contains the data types, permitted values, and any relationships or conditional requirements between characteristics of a reported power analysis. The JSON schema serves as a formal machine-readable specification of the reporting standard. Researchers are not expected to report power analyses in this format, but to report power analyses clearly enough such that software can parse their writing, and add it to a schema automatically (without the use of large language models). An example of the full .JSON file is presented below, and is followed by a detailed explanation of each aspect of a power analysis that should be provided for full reporting.
-->


### Unique identifier

<!--
LDB: This is mainly relevant to a schema:

We propose that each power analysis should be assigned a unique identifier. This identifier can be used to create an explicit link between the research question, the statistical test and criteria used to address it, and the corresponding power analysis. 
-->
Power analyses are often reported with insufficient detail to determine which hypothesis or statistical test they were intended to support [@thibault_2024; @bakker_2020; @mesquida_2025]. This issue is particularly relevant in studies that include multiple hypotheses or multiple statistical tests. A complete report must allow the power analysis to be linked to the corresponding hypothesis and statistical test. For tests that produce more than one result, such as a factorial ANOVA or regression with multiple predictors, it must also be clear what effect the power analysis is for; the power for a main effect may not be the same as the power for an interaction.

<!--
Move to discussion?:

Combined with hypotheses and statistical tests also in a machine-readable format [@lakens_2021], such links could facilitate automated checks of whether all primary analyses are supported by a corresponding power analysis. Where researchers often list and number hypothesis (e.g., H1, H2, H3), it should be just as normal to number multiple tests and power analyses.
-->

### Target of power analysis

Specific terminology around power analyses can be complex and confusing. An *a priori* power analysis calculates sample size before data collection, given desired power, a target effect size, the alpha criterion, and specific test parameters. A *compromise* power analysis calculates power and alpha from sample size and effect size, given a specific ratio between Type 1 (alpha) and Type 2 (1 - power) error rates. A *criterion* power analysis calculates the alpha required for the desired power, given a specific sample size and effect size. A *sensitivity* power analysis calculates the smallest effect size that can be detected with desired power given a sample size and alpha criterion. 

After data collection, researchers may also calculate the hypothetical power of a study design based on the final sample size and an assumed effect size of interest, often termed a *post hoc* power analysis. This provides information about the sensitivity of the study design given the sample size and can be understood as the post-data-collection counterpart of an *a priori* power analysis. However, the term *post hoc* power analysis is frequently used ambiguously. It may alternatively refer to calculating power using the effect size observed in the data, which provides no additional information beyond the *p*-value [@yuan_2005; @lakens2022]. Researchers have therefore recommended avoiding calculating power based on observed effect size [@hoenig_2001; @yuan_2005; @okeefe_2007; @lakens2022]. 

<!--
For clarity, we propose using *effect size sensitivity* for calculating the smallest effect size that can be detected with desired power, *power sensitivity* for calculating power given a hypothetical effect size of interest, and *observed power* for calculating power using the effect size observed in the data.
-->

Due to the confusion around terminology, we propose a straightforward classification of power analyses based on the target being calculated: *power*, *sample size*, *effect size*, or *alpha*. Further distinctions between post hoc and observed power analysis can be determined from the effect size justification (detailed below).

### Power

Although a power of 0.8 is commonly used as a default, researchers should explicitly report their chosen level of power, and there are good reasons to increase the statistical power 0.9 or 0.95 where possible. <!-- cite? -->  When power is an input to the calculation, it is useful to include a justification for the value used. To facilitate machine readability and automated extraction, power should be reported as a numeric value between 0 and 1 (e.g., 0.80 rather than 80%).

### Sample size

The sample size used in the power analysis should be reported explicitly. Researchers should both report the total sample size, and where applicable, the sample size per group. Reporting the total and per group sample size allows the power analysis to be reproduced. For studies with multiple observation per individual, information about the number of observations should be provided as well. This number should also be justified if it is an input to the power analysis [@lakens2022].

### Effect size

The effect size metric (e.g., Cohen's $d$, $r$, odds ratio) and magnitude (e.g., 0.5, 1.3) used in the power analysis should be reported. However, a common issue in the reporting of power analyses is that neither the effect size metric nor its magnitude are reported [@thibault_2024; @bakker_2020; @mesquida_2025]. Reporting only the numerical value (e.g., an effect size of 0.1) is insufficient because the same value can represent different quantities depending on the metric used. For example, an effect size of 0.1 could refer to Cohen's $d$, Cohen's $f$, or $\eta_p^2$ ). Reporting the metric allows readers to determine whether the assumed effect size corresponds to the effect of interest, and increases the reproducibility of the power analysis. Reporting the exact magnitude, rather than t-shirt sizes like "small", "medium", or "large", allows readers to evaluate whether the assumed effect size is plausible, justified, and appropriate for the research question.

Reporting the metric may not always be sufficient trace back what it was based on. Researchers may convert the effect size they use in a power analysis from an effect size reported in another metric, calculate an effect size from summary statistics, or combine multiple effect sizes in the literature to obtain the value used in the power analysis. For example, researchers may combine information from multiple studies by averaging reported effect sizes, or apply bias-adjustment methods to an effect size reported in a previous study or meta-analysis. Although this recommendation goes beyond simply reporting the power analysis, it is essential information to be able to evaluate the validity of a power analysis. Researchers should transparently report how the effect size used in the power analysis was obtained or calculated. For example, if an effect size was converted from another metric, researchers should report the original metric and the conversion formula used. If the effect size was calculated by the researchers, the calculation procedure should be described. This information allows readers to evaluate the assumptions underlying the selected effect size and improves the reproducibility of the power analysis.

The source of the effect size is important for interpreting a power analysis. There are six common approaches that researchers use to justify the effect size of interest in a power analysis: effect sizes from previous studies, meta-analyses, pilot studies, Cohen's *d* thresholds, field-specific thresholds, and the smallest effect size of interest (SESOI) [@lakens2022]. Reporting this information again goes beyond the statistical properties of a power analysis, but is essential when evaluate whether the sample size justification is valid. We therefore think that in addition to reporting how the effect size was determined, researchers should provide a justification for why this effect size was used in the power analysis.

This information is especially important to be able to detect common problems with the choice of an effect size of interest in *a priori* power analyses. When effect sizes are derived from previous studies or meta-analyses, researchers should explain how the current study is sufficiently similar to the planned study, as studies might differ in the population, study design, manipulation, intervention, or even outcome measures. Furthermore, researchers should reflect on the (possibly large) uncertainty in the effect size estimate of the original study. This is also true if the effect size comes from a pilot study that the researchers performed themselves.

Instead of using an empirically observed effect size for the power analysis, researchers may rely on Cohen's *d* thresholds for small, medium, and large effects, or field-specific benchmarks. This practice can be problematic because it uses general benchmarks to design a specific study that is likely not similar to the general study in a field. Consequently, choosing an effect size solely on the basis of Cohen's *d* thresholds or other field-specific benchmarks may overlook the contextual factors that determine whether the effect size is appropriate for the planned study [@primbs_2022].

A better approach is to justify the effect size based on a smallest effect size of interest (SESOI), which specifies the effect a researcher does not want to miss, either because it is considered theoretically or practically meaningful for the research question. Regardless of the justification for the effect size is used to compute power, it should not simply be stated, but researchers should justify why it is a relevant effect size to design the study around.

### Alpha criterion

An important component of a power analysis is the decision criterion, usually a critical alpha. Although 0.05 is commonly used as a default alpha level in hypothesis tests, the alpha level should be explicitly reported because it determines the threshold for statistical significance and directly affects the power of a test. Reporting the alpha level also allows peers to evaluate whether it matches the alpha level used in the study. For example, a study may adjust the alpha level to account for multiple comparisons in the results section, but use a different alpha level in the power analysis. As corrections for multiple comparisons lower the alpha level, they also lower the statistical power of the test. To facilitate machine readability and automated extraction, the alpha level should be reported as a numeric value between 0 and 1 (e.g., 0.05 rather than 5%). Researchers should also provide a justification explaining the rationale for this choice [@lakens_justify_2018], such as adjustments for multiple comparisons. 

The decision to reject the null hypothesis may rest on more than one criterion, or on criteria that are not only the comparison of one p-value to an alpha level. When this decision is more complex, power analysis will usually need to be conducted by simulation and the specific decision criteria must be identical between the power analysis and the analysis reported in the results. 

### Test parameters

Researchers should provide all information about the statistical test(s) that will affect the power analysis. For example, power analysis software such as G\*Power requires users to specify the test family (e.g., *t* tests, *F* tests) and specific statistical test (e.g., Wilcoxon signed-rank test for matched pairs, repeated measures ANOVA). Ensure that this is the same test that will be used to test your hypothesis.

<!--
Specifying the type of effect clarifies which hypothesis the power analysis was designed to evaluate and allows readers to determine whether the power analysis corresponds to the intended research question.

LDB: I feel like this shouldn't be the way we determine the match with a hypothesis! Tests should explicitly link to hypotheses (we argue in lakens_2021) and then power analyses explicitly link to test-effect or to hypotheses? In the MRHT paper, we create decision criteria for integrating the results of 1 or more tests to support, refute, or not-support a specific hypothesis -- that is the target of a power analysis. Which only corresponds to 1 effect in 1 test for simple things. 
-->

Most statistical tests require additional parameters beyond power, sample size, effect size and alpha to fully specify the power analysis. These test parameters depend on the test family and the design configuration. These parameters often have default values, but it is essential to specify these explicitly to facilitate replication with different software. For correlations and *t*-tests, the directionality of the hypothesis should be reported using the tail parameter. A repeated measures ANOVA with a within-between interaction requires the number of groups, the number of measurements, a non-sphericity correction, and the correlation among repeated measures. This correlation represents the expected association between repeated observations and influences the variance estimate used in the power calculation. In complex designs, this would require specifying a correlation matrix. The more complex a test is, the more details of a test need to be specified, up to a point where it might be easier to share the code for the power analysis than to describe all details of the test verbally.

### Software

The software used to conduct the power analysis should be reported to allow peers to reproduce the analysis using the exact same software, and where relevant, any add-ons used (e.g., the specific R package). Different software packages and even different versions of the same software can return slightly different calculations for the same settings, so also report software and package versions. Researchers should also provide a link to the script, notebook, or other file that contains the code used to perform the power analysis. The link is preferably a Digital Object Identifier (DOI) that links to a file in a stable data repository. If the software used is point-and-click, researchers can link to a screenshot or a detailed written description of all settings.

## Reporting

One reason power analyses lack key information is the absence of specific reporting guidelines that editors, reviewers, and researchers can follow. For example, the American Psychological Association's (APA) *Journal Article Reporting Standards* (JARS) provides reporting standards for several aspects of scientific articles [@appelbaum2018], such as p-values and confidence intervals, but it provides no guidance on how to report power analyses. Furthermore, JARS reporting guidelines specify *what* information to report, but rarely provide detailed information about *how* the information should be presented. For example, JARS for quantitative articles state that researchers should describe the "power analysis, or methods used to determine precision of parameter estimates" but does not specify how this information should be reported. Given the importance of power analyses for the evaluation of research findings, a reporting standard that specifies both the required information and a consistent structure for how this information should be presented could substantially improve current research practices.

```{r f_test_example}
f_statistic <- 5.23
numerator_df <- 1
denominator_df <- 198
n <- 200

f_test <- effectsize::F_to_eta2(
  f = f_statistic, 
  df = numerator_df, 
  df_error = denominator_df, 
  ci = 0.90, 
  alternative = "two.sided"
)

p_value <- pf(
  q = f_statistic, 
  df1 = numerator_df, 
  df2 = denominator_df, 
  lower.tail = FALSE
)

f_test_report <- sprintf(
  "*F*(%d, %d) = %.2f, *p* %s %.3f, $eta_p^2$ = %.2f, 90%% CI = [%.2f; %.2f]",
  numerator_df,
  denominator_df,
  f_statistic,
  ifelse(p_value < .001, "<", "="),
  ifelse(p_value < .001, .001, p_value),
  f_test$Eta2_partial,
  f_test$CI_low,
  f_test$CI_high
)
```

A useful precedent exists in the standardised reporting of statistical tests following APA guidelines. Researchers are familiar with reporting statistical tests such as *t*-tests and *F*-tests in a consistent order that includes the test statistic, degrees of freedom, *p*-value, effect size, and its confidence interval. For example, an *F*-test may be reported as (`r f_test_report`). These standardised reporting convention allows readers to quickly identify and interpret the information they convey. At the same time, standardised reporting practices have the important benefit that they enable automated extraction of tests, which is used by tools such as Statcheck to reliably parse statistical tests and verify reported *p*-values [@nuijten_2015].

The same approach can be applied to power analyses. The standardised reporting of power analyses should ensure that information is interpretable by human readers but also structured in a way that automated tools like Metacheck <!-- cite --> can extract using deterministic approaches (e.g., regular expressions). The extracted information can then be put in a structured format, which can be validated against a schema to ensure that all required fields are present and that reported values conform to the specified constraints. <!-- I made this more generic because JSON is jargon-y and not the only possibility, just a standard we like -->

As explained above, a fully reported power analysis requires the type of power analysis, power, the sample size, the effect size specified with a clear metric and justification, the critical alpha, relevant details of the statistical test, and the software used. G*Power's "Protocol of power analyses" contains the relevant input and output parameters to report. 

:::{.callout-note}
## Example G*Power protocol output

```
F tests - ANOVA: Repeated measures, within-between interaction

Analysis:	A priori: Compute required sample size 
Input:  Effect size f              = .12
        α err prob                 = 0.05
        Power (1-β err prob)       = .9
        Number of groups           = 2
        Number of measurements     = 2
        Corr among rep measures    = 0.5
        Nonsphericity correction ε = 1
Output:	Noncentrality parameter λ  = 10.7136000
        Critical F                 = 3.8924944
        Numerator df               = 1.0000000
        Denominator df             = 184
        Total sample size          = 186
        Actual power               = 0.9025241
```
:::


Combining these properties in our proposed standard report for power analyses could lead to a sentence such as:

> We determined sample size using G\*Power 3.1.9.6 (Mac) [@gpower] for "ANOVA: Repeated measures, within-between interaction" with two groups and two measurements, using a critical alpha of 0.05 and assuming a correlation among repeated measures of 0.5 and no nonsphericity correction (ε = 1). The effect size of interest was f = 0.12, based on the previously published study by Someone et al. (2025). With a desired power of 0.90, this returned a sample size of 186 observations.


Alternatively, and just as statistical tests are abbreviated instead of communicated in prose, one could write:

> A power analysis for an effect size of interest determined by a previous study (Someone et al., 2025) indicated that 186 observations should be collected (sample size power analysis: power = 0.80; N = 186; f = 0.12; α = 0.05; F-test, A:B of A(2w)\*B(2b), r = 0.5, ε = 1; G\*Power, 3.1.9.6 (Mac), [@gpower]).

Generically, the format is ordered as follows: Target of power analysis, power, sample size, effect size, critical alpha, test details, software. For example, a sensitivity power analysis based on the pwr package in R would be reported as:

> A power analysis indicates that effects of d = 0.69 could be detected with the desired 90% power (effect size power analysis: power = .90; N = 90; d = 0.69; α = 0.05; t-test, type = "two.sample", alternative = "greater"; pwr package, @R-pwr).

As tests get more complicated and require more parameters, it may become necessary to reference an external file for all of the details. Summarise any parameters that a reader may use to interpret the power analysis, and include the full specification in the referenced script.

> We conducted a power analysis to determine that a between-group difference of 1.2 points was the smallest effect size we had 80% power to detect with 120 subjects and 50 stimuli (effect size power analysis: power = .80; N(subj) = 120, N(stim) = 50; effect size = 1.2 points; α = 0.01; linear mixed effects model, points ~ group + (1 | subj_id) + (1 | stim_id), groups = 2; simulation in R with faux, @R-base, @R-faux, https://doi.org/example). The power and alpha were chosen due to journal policy requirements; the number of subjects and stimuli were chosen due to availability constraints.

In fields where a specific complex analysis is the norm, and certain parameters are essential for interpretation, researchers should develop norms around what parameters should be communicated in text and how. 

## Automated checks

One way to improve the reporting of power analyses is by developing tools that perform automated checks and provide feedback to researchers about how to improve their reporting. Metacheck [@R-metacheck], a suite of tools for checking research outputs for best practices, has one such tool in the built-in `power` module. This module identifies paragraphs that are likely to contain a power analysis using text-matching and returns the paragraphs along with advice on the components of a complete power analysis in a report.

Extracting parameters of a power analysis from unstructured scientific text is challenging because the same information can be phrased in many different ways. For example, the sample size could be communicated as "power analysis calculated 65 participants per group", "One hundred and thirty was the sample size", or "We used power calcuation to determine the sample size (N = 130)".  Consequently, deterministic rule-based approaches fail to identify parameter information that is present but expressed in an unexpected way. To address this limitation, the `power` module can optionally use an LLM to interpret flexible human language and provide even more targeted advice.

```{r}
#| label: llm_power
#| eval: FALSE ## Set to TRUE to rerun module

# Opt-in LLM
llm_use(TRUE)

# Choose model
llm_model("ollama/qwen3.5:9b")      

# Increase the number of calls. Maximum number set to 30
llm_max_calls(2500)

# Run power module
res_power_llm <- module_run(psych_science, "power")

# Save the results
save(res_power_llm, file = "data/res_power_llm.RData")
```

```{r}
# Load data
load("data/res_power_llm.RData")

# Number of power analyses
n_power_analyses <- nrow(res_power_llm$table)
```

To illustrate this approach, we applied the LLM-version of the `power` module to a sample of `r length(psych_science)` open-access articles from *Psychological Science*. @tbl-power_llm shows the proportion of missing components from `r n_power_analyses` power analyses. Although the LLM-based approach can identify information expressed in a wider range of ways than rule-based methods, some parameters remain missing because they were not reported or could not be reliably extracted from the text.

<!--#  Below we show the prevalence of missing information using the LLM. This was already reported in the Metacheck paper. Instead of reporting the same information twice perhaps we can show an example of poorly reported power analysis?-->

```{r}
#| label: tbl-power_llm
#| tbl-cap: !expr "sprintf('Proportion of missing components from %d power analyses.', n_power_analyses)"

# Estimate the number and proportion of components from power analyses that are missing
tbl_power_llm <- res_power_llm$table |> 
  select(software_other:power_type) |> 
  mutate(
     # Combine software fields 
    software_combined = coalesce(
      software, 
      software_other
      ),
    # Combine effect size metric fields
    effect_size_metric_combined = coalesce(
      effect_size_metric,
      effect_size_metric_other
    ),
    # Combine statistical test fields
    statistical_test_combined = coalesce(
      statistical_test_other,
      statistical_test 
    )
  ) |> 
  select(
    -c(
    software, 
    software_other, 
    effect_size_metric, 
    effect_size_metric_other,
    statistical_test_other,
    statistical_test
    )
  ) |>
  mutate(
    across(
      -power_type,
      ~ is.na(.x)
    ),
    power_type = power_type == "unknown"
  ) |>
  summarise(
    across(everything(),
           list(
             n = sum,
             pcnt = mean
           )
    )
  ) |> pivot_longer(
    cols = everything()
  ) |>
  separate(
    name, 
    c("Component", "name"), 
    sep = "_(?!.*_)"
    ) |>
  pivot_wider() |>
  mutate(pcnt = sprintf("%d%%", round(pcnt*100)))

# Display the table
comp_ord <- c("power_type", 
              "power", 
              "sample_size", 
              "effect_size", 
              "effect_size_metric_combined",
              "alpha_level",
              "statistical_test_combined",
              "software_combined")
tbl_power_llm |> 
  mutate(Component = factor(Component, comp_ord)) |>
  arrange(Component) |>
  mutate(
    Component = gsub("_combined", "", Component),
    Component = gsub("_", " ", Component)
  ) |>
  flextable() |>
  align(
    align = "right", 
    j = "pcnt", 
    part = "all"
    ) |>
  set_header_labels(
    n = "Missing",
    pcnt = "Percent"
    ) |>
  width(
    1:3, 
    c(2, 1, 1)
    )
```

Although LLMs provide a flexible way to parse natural language variation, they treat the symptoms of inconsistent reporting rather than its root cause. When information is inherently ambiguous or incomplete, an LLM may misclassify a power analysis or extract incorrect parameters. Moreover, using LLMs for routine information-extraction tasks incurs substantial computational and energy costs. A preferable long-term solution is to reduce the need for LLM-based approaches by making scientific text easier to extract in the first place. This goal is best achieved through standardised reporting conventions, such as those proposed in this paper.

<!--
LDB: probably we don't need this paragraph now?
LDB: Is machine-readability a property of this first goal or the second?

To improve the way power analyses are reported in the scientific literature, we propose a two-step solution. First, we will develop standardised reporting guidelines that provide clear and universal guidance on how to transparently and completely report power analyses. Importantly, these guidelines will be structured in a format that facilitates machine readability, so that reporting standards can be consistently interpreted and applied by automated tools without relying on large language models to interpret the information. Second, we will formalise these reporting guidelines in a structured schema that specifies the information that should be reported for each component of a power analysis. This schema will be used in the Metacheck power module to validate the information reported in power analysis statements. This will facilitate the identification of missing or incorrectly reported information and provide researchers with guidance on best reporting practices.
-->

:::{.callout-important}

Describe the future of machine readable power analyses using scienceverse here. What do we need to develop and how can it all work together?

- Each hypothesis, analysis, dataset, and power analysis has an ID
- Hypotheses contain specific criteria (p < .05; M1 < M2, ect.) that link to specific analysis outputs, and decision rules (criterion 1 *or* criterion 2)
- Datasets can contain a data-generating model to describe how to simulate data with the expected effect(s)
- The power analysis object links a hypothesis, which contains criteria that reference analyses. It contains the essential info like power, sample size, effect size, alpha
- This can all automatically generate power by using the DGM to simulate data, the analyses to analyse the simulated data, and the criteria/decision rules to determine if the null hypothesis should be rejected. This allows for complex analyses that don't have an analytic power solution, and also complex combinations of decision criteria.  
- For simpler analytic power analyses, the power analysis object can contain (pseudo)code for pwr or G*Power sufficient to replicate the analysis, otherwise link to a script for complex analyses.

:::


## Conclusion

Standardizing the way information is reported in scientific articles has benefits beyond improving the transparency of individual power analyses. When information is consistently structured across articles, automated tools can more reliably extract and organize reported information.

The proposed standardized reporting represents an initial step toward a more transparent and clear representation of power analyses by imposing a more consistent structure on the information required to evaluate and reproduce power analyses, while also facilitating information extraction using rule-based automated tools.

The current version focuses on commonly used statistical tests, including correlations, *t*-tests, and *F*-tests. Extending the standard to accommodate more complex analytic approaches, such as power analyses for mixed-effects models or simulation-based power analyses, will require identifying additional parameters and reporting requirements. Developing a comprehensive standard that covers the diversity of power analysis approaches used across disciplines will therefore require a community effort involving researchers and other stakeholders, such as journals. Such collaborative development can ensure that future extensions of the schema reflect the needs of different research communities while maintaining a consistent structure that supports both human interpretation and automated extraction. Future developments could also explore workflows in which software generates machine-readable representations directly, reducing the need for manual data entry while preserving the advantages of structured reporting.

```{r}
#| label: session-info
#| include: false

capture.output(
  sessioninfo::session_info(),
  file = "session-info.txt"
)
```

## References
